% Documentation of the NStart sensor

\subsection{\texttt{Sensor.NStart}}

\paragraph{Description}

This plugin injects possible translation starts as predicted by the
\textsc{NetStart} program. The plugin reads the prediction of the
programs from two files whose names are derived from the sequence name
by adding the \texttt{.starts} and \texttt{.startsR} suffix
(respectively prediction for the forward and reverse strand).

The sensor is activated by setting the value TRUE for the parameter
\texttt{Sensor.NStart.use} in the parameter file. The score for acceptor
and donor prediction is rescaled by the parameters {\tt NStart.startP} and
{\tt NStart.startB} (see below).

Here is an example of NG2 parameters definition.
\begin{Verbatim}[fontsize=\small]
NStart.startP*     0.052
NStart.startB*     0.308
Sensor.NStart.use  TRUE    # Use NStart sensor
Sensor.NStart      1       # Sensor priority
\end{Verbatim}

\paragraph{Input files format}

The files with \texttt{.starts} and \texttt{.startsR} suffix are
obtained by running NetStart which can be obtained at
\texttt{http://www.cbs.dtu.dk/services/NetStart/} and using the
detailed output of the software.

Here is an extract from \texttt{SYNO\_ARATH.fasta.starts}:
\begin{Verbatim}[fontsize=\small]
  1089    0.256     -
  1146    0.214     -
  1251    0.618     Yes
  1299    0.197     -
  1474    0.526     Yes
  1535    0.112     -
  1559    0.490     -
  1638    0.401     -
  1674    0.569     Yes
  1678    0.147     -
  1740    0.299     -
  1752    0.187     -
[...]
\end{Verbatim}

To run NetStart, the following parameters are used (for Arabidopsis
thaliana): \texttt{netstart -at <Fasta sequence>}.

\paragraph{Filtering input information}

No filtering.

\paragraph{Integration of information}

The integrated score for start prediction is read (column 2).  The
score read $s$ is rescaled using the {\tt NStart.startP} and {\tt
NStart.startB} as follows:

\[s' = B*s^P\]

All predictions that use a predicted start receive a $\log(s')$
penalty while those that go through a predicted start while they
could have used it receive a $\log(1-s')$ penalty.


\paragraph{Post analyse}

No post analyse.

\paragraph{Graph}

Predicted starts are visible on exonix tracks as blue vertical lines
whose length indicates the site score.





