% Documentation of the ProStart sensor

\subsubsection{\texttt{Sensor.ProStart}}
\label{prostart}
\paragraph{Description}

This plugin is dedicated to inject possible translation starts in prokaryote sequences.

This sensor is activated by setting the value 1 for the parameter \texttt{Sensor.ProStart.use} in the parameter file.

\texttt{ProStart.stackFile} and \texttt{ProStart.loopFile} : contain the energetic parameters for computing the RBS 
(Shine-Delgarno) hybridization energy. Taken and adapted from Turners website.
\EuGene\ use a simplified RNA binding model which ignores stabilizing loops (eg. tetraloops). 
The algorithm is essentially a free end-gap global alignement algorithm that takes into account bulges, loops and helices.

\texttt{ProStart.alpha*} and \texttt{ProStart.beta*}: Start parameters $\alpha$ and $\beta$. 
These parameters are used to control how the RBS quality affect the probability of use of a given START codon. 
If $E$ is the hybridization energy of the RBS, the following probability is used:

\[\frac{\alpha}{1+\beta\cdot\exp(E/RT)}\]
  
So, $\alpha$ gives a basic probability for a START to be used. For $\beta$, the
higher it is, the stronger the influence of the RBS quality on the use of
the corresponding START.  START probability is also affected by their type: 
\texttt{atg} is the preferred START codon, then \texttt{gtg} and \texttt{ttg}.

Here is an example of ProStart parameters definition

\begin{Verbatim}[fontsize=\small]
ProStart.matchlen 14
ProStart.matchoffset 3
ProStart.RBSPattern attcctcca # E coli (16S genetics/ Steitz J.A.)
ProStart.alpha* 0.4
ProStart.beta* 6.0
ProStart.stackFile stack.dat
ProStart.loopFile loop.dat

Sensor.ProStart.use 1
Sensor.ProStart 10
\end{Verbatim}
                                                           
\paragraph{Input files format} No input files needed

\paragraph{Integration of information}

If $E$ is the hybridization energy of the RBS, the following probability is used:

  \[p=\frac{\alpha}{1+\beta\cdot\exp(E/RT)}\]

All predictions that use a predicted start receive a $\log(p)$
penalty while those that go through a predicted start while they
could have used it receive a $\log(1-p)$ penalty.

\paragraph{Post analyse}

No post analyse.

\paragraph{Graph}

Predicted starts are visible on exonix tracks as blue vertical lines
whose length indicates the site score.
